---
title: "Own Account Agriculture Data Request"
author:
  - name: "Renato Vargas"
    id: rv
    email: renovargas@gmail.com
    affiliation: 
      - name: Consultant for The World Bank
            
format:
  html:
    toc: true
    number-sections: true
    number-depth: 3
    highlight-style: github
  # docx:
  #   toc: true
  #   number-sections: true
  #   highlight-style: arrow
  # pdf:
  #   toc: true
  #   number-sections: true
  #   colorlinks: true
editor: source
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: apa-6th-edition.csl
---

# Introduction

In this data request, we aim to describe the people who receive some income from agriculture (or produce subsistence agriculture) but are not classified as agriculture workers. Relevant aspects to consider are:

- Age, 
- Other sources of income (maybe they have another job?), 
- Position in the income distribution (for example the decile they’re in).
- Other relevant households characteristics.


## Preamble

We start with a clean environment, making sure that any objects from a previous session are not present and declare some relevant variables.

```{r}
#| warning: false
#| messAge: false

# Clean workspace
rm(list = ls())

# Georgia country ISO code
iso <- "GEO"

# Survey year
survey_year <- 2023

# Poverty line
pline <- 212.8149

# Exchange rate USD per GEL (Multiply by this to show results in USD)
er <- 0.37
```

Rather than calling our libraries as we go, we will make sure we have everything we need from the beginning.

```{r}
#| output: false
#| lst-label: lst-load-packAges

library(tidyverse) # includes dplyr, ggplot2, purr...
library(haven)     # to read SPSS and Stata datasets
library(readxl)    # to read from MS-Excel
library(openxlsx2)  # to write to MS-Excel.
library(gt)        # pretty tables
library(flextable) # pretty tables
# library(Hmisc)   # Estimates deciles, quintiles but use :: version too
# library(ineq)    # Inequality measures
# library(acid)    # Inequality measures we use acid::weighted.gini()

# GeopackAges
library(sf)        # to read and write shapefile maps
library(terra)     # to perform geocalculations
library(tmap)      # for static and interactive maps
```

# Datasets

We then load the datasets that we need for this study. These are based on Georgia's Integrated Living Conditions Survey 2023 [@geostat_integrated_2023]. We make a note that the household identification variable is `UID`.

```{r}
#| warning: false
#| messAge: false
#| output: false
#| lst-label: original-datasets

## Household Unique ID, Weights, Location and other basic variables
hh_basics <- read_sav(
  "data/ilcs_2023/sysschedule.sav") |>
  mutate(
    UID = as.integer(UID))

# Household size (includes no. of family members)
hh_size <- read_sav(
  "data/ilcs_2023/familysize.sav")|> 
  mutate(
    UID = as.integer(UID))

# Processed income at household level
hh_income <- read_sav(
  "data/ilcs_2023/tblincomes.sav")|> 
  mutate(
    UID = as.integer(UID))

# Consumption aggregate at household level 
hh_expenditure <- read_sav(
  "data/ilcs_2023/tblexpenditures.sav")|> 
  rename(# rename total expenditure variables
         total_expenditure = MTlianixarjebi_,
         total_expenditure_aeq06 = MTlianimoxmareba_EqAdScale,
         total_expenditure_aeq08 = Mtlianimoxmareba_EqAdScale_08) |> 
  mutate(
    UID = as.integer(UID))

# Characteristics of the dwelling
hh_chars <- read_sav(
  "data/ilcs_2023/tblshinda01.sav")|>
  mutate(
    UID = as.integer(UID))

# Persons (pp)
pp <- read_sav(
  "data/ilcs_2023/tblshinda02.sav") |> 
  mutate(
    UID = as.integer(UID),
    MemberNo = as.integer(MemberNo))

# Labor (pp)
pp_labor <- read_sav(
  "data/ilcs_2023/tblshinda05_1.sav") |> 
  mutate(
    UID = as.integer(UID),
    MemberNo = as.integer(MemberNo),
    Q5  = as.integer(Q5),
    Q12 = as.integer(Q12)) 

# Poverty
poverty <- read_dta(
  "data/ilcs_2023/POVERTY_stata.dta") |> 
  rename(official_pline = pline) |> 
  mutate(
    UID = as.integer(UID),
    poor_2023 = if_else(aecons < pline, 1, 0))

# Ind. Poverty
ind_poverty <- read_dta(
  "data/ilcs_2023/IND_POVERTY_stata.dta") |> 
  rename(MemberNo = memberno) |> 
  mutate(
    UID = as.integer(UID),
    MemberNo = as.integer(MemberNo))

# Food diary
food_q <- read_sav(
  "data/ilcs_2023/tblconsumption.sav") |> 
  rename(UID = UID)

food_price <- read_sav( 
  "data/ilcs_2023/tblavgprices.sav")

# Maps
adm1 <- sf::read_sf("data/gis/geo-adm1.shp") |>
  dplyr::select(RegNo, region, ADM1_PCODE, ADM1_EN, ADM1_KA, geometry) |>
  dplyr::arrange(ADM1_PCODE)

regions <- as.data.frame(adm1) |> 
  select(-geometry)
```

We also need look-up tables.

```{r}
#| warning: false
#| messAge: false
#| label: look-up-tables

sam_activities <- read_excel(
    "data/sam/classifications.xlsx",
    sheet = "SAM-REV2",
    col_names = T,
    col_types = c("text", "text", "text","text", "numeric")
  )

sam_factors <- read_excel(
    "data/sam/classifications.xlsx",
    sheet = "SAM factors",
    col_names = T,
    col_types = "text",
  )

coicop <- read_excel(
    "data/sam/classifications.xlsx",
    sheet = "COICOP",
    col_names = T,
    col_types = "text",
  ) |> 
  mutate(simple_code = as.integer(gsub("\\.", "", Coicop)))

coicop_filtered <- coicop |> 
  filter( nchar(as.character(simple_code)) >= 5)

```


We also have Continuous Labor Survey data at the individual level, which will come in handy if we do not get access to the labor part of the ILCS. See data folder for documents describing the datasets.

```{r}
#| warning: false
#| messAge: false
#| output: false
#| lst-label: labor-survey

# Labor Force Survey
lfs_2023 <- read_sav(
  "data/lfs_2023/LFS_ECSTAT_ENG_2023.sav") |> 
  rename(UID = UID)

# Labor Force Survey Demographic Characteristics
lfs_2023_dem <- read_sav(
  "data/lfs_2023/LFS_Demographic_ENG_2023.sav") |> 
  rename(UID = UID)

```

# Generating descriptor variables (SAM disaggregates)

## Skill level

For skill level, we will use information on schooling from `pp$Education` (`TblShinda02`), which has the following levels:

1.  Illiterate
2.  Do not have primary education but can read and write
3.  Pre-primary education
4.  Primary education
5.  Lower secondary education
6.  Upper secondary education
7.  Vocational education without secondary general education
8.  Vocational education on the base of lower secondary education with secondary general education certificate
9.  Vocational education on the base of secondary general education (except higher professional education)
10. Higher professional program
11. Bachelor or equivalent
12. Master or equivalent
13. Doctor or equivalent

We need three skill levels for our SAM template, so we map these levels to:

Low skill (1 - 5): Illiterate through lower secondary.
Medium skill (6 - 9): Upper secondary through vocational education.
High skill (10 - 13): Higher professional program through Doctor.


```{r}
#| warning: false
#| messAge: false
#| label: skill-level

pp_factor_descriptors <- pp |>
  select(UID, MemberNo, Gender, Age, Education) |> 
  mutate(
    MemberId = 
      paste0(sprintf("%06d", UID), sprintf("%02d", MemberNo))) |>
  mutate(Gender = factor(
    Gender,
    levels = c(1, 2),
    labels = c("Female", "Male")
  )) |> 
  mutate(
    SkillLevel = case_when(
      Education >= 0 & Education <= 5 ~ 1,
      Education > 5 & Education <= 9 ~ 2,
      Education > 9 & Education <= 13 ~ 3,
      TRUE ~ NA ) ) |> 
  mutate(
    SkillLevel = factor(
      SkillLevel, 
      levels = c( 1, 2, 3),
      labels = c( "Low Skill", "Medium Skill", "High Skill"))
  )
```

Now that we have skill levels, we need to add information on urban/rural (from `hh_basics`) and quintile (from `ind_poverty`), and type of income earner (from `pp_labor`).

```{r}
#| warning: false
#| messAge: false
#| label: factor-income-descriptors

urb_rur <- hh_basics |> 
  select(UID,QuartNo, UrbanOrRural, RegNo, Weights) |> 
  mutate(
    UrbanOrRural = factor(
      UrbanOrRural,
      levels = c(2,1),
      labels = c("Rural", "Urban")
    ),
    weight = Weights / 4
  )

quintiles <- poverty |> 
  select(UID, quintilc) |> 
  rename(Quintile = quintilc) |> 
  mutate(
    Quintile = factor(
      Quintile,
      levels = c(1:5),
      labels = c("Q1", "Q2", "Q3", "Q4", "Q5")
    )
  )

income <- hh_income |> 
  select(
    UID, 
    ShemDaq, # labor income
    ShemTviTdasaqm, # self_employment_income
    Shem_Sof, # agriculture income
    Qonebidan, # property income
    PensStipDaxm, # Pensions, stipends
    Ucxoetidan, # Remittances from abroad
    Axloblebisagan, # Money received as a gift
    ShemosavaliDaTransf # Total income
    ) |> 
  rename(
    labinc = ShemDaq,
    selfemp = ShemTviTdasaqm,
    aginc = Shem_Sof,
    propertinc = Qonebidan,
    pensions = PensStipDaxm,
    remittances = Ucxoetidan,
    gifts = Axloblebisagan,
    totalinc = ShemosavaliDaTransf
  )

pp_factor_descriptors <- pp_factor_descriptors |> 
  left_join(urb_rur, join_by(UID)) |> 
  left_join(quintiles, join_by(UID)) |> 
  left_join(income, join_by(UID))
```

## Labor status and Economic Activities

We work with labor status from `Shinda05_1`. Since, upon import NACE 2 codes are converted to numbers, we need to convert them back to text, so that we can keep zeros to the left for proper order. We then extract the first two digits and find the correspondence to Rev. 2 from the SAM using the look-up table `sam_activities`. For proper order, we convert the SAM activities columns for job 1 and job 2 to factor, using the order from the dataset `sam_factors`.

```{r}
#| warning: false
#| messAge: false
#| label: nace-codes

pp_emstatus <- pp_labor |> 
  mutate(
    MemberId = 
      paste0(sprintf("%06d", UID), sprintf("%02d", MemberNo))) |>
  mutate(
    # Job 1 NACE Rev 2 code. 
    Q5  = if_else(!is.na(Q5),paste0(sprintf("%04d", Q5)), NA),
    # Job 2 NACE Rev 2 code.
    Q12 = if_else(!is.na(Q12),paste0(sprintf("%04d", Q12)), NA)) |> 
  mutate(
    job1 = if_else(!is.na(Q5),substr(Q5, 1, 2), NA),
    job2 = if_else(!is.na(Q12),substr(Q12, 1, 2), NA)
  ) |> 
  # We match to Rev 2 and SAM classifications (for job 1 and job 2)
  left_join(
    sam_activities[,c(1,3)], 
    join_by(job1 == rev2_2d)) |> 
  left_join(
    sam_activities[,c(1,3)],
    join_by(job2 == rev2_2d),
    suffix = 
      c("_job1", "_job2")) |> 
  # And convert to factors for proper order
  mutate(
    SAM_job1 = factor(
      SAM_job1, 
      levels = sam_factors$SAM,
      # labels = sam_factors$SAM_description
      ),
    SAM_job2 = factor(
      SAM_job2, 
      levels = sam_factors$SAM,
      # labels = sam_factors$SAM_description
      )
  )
```

## Types of income

Before making our multi-dimensional tables, we need to identify different types of income. f-lab (wAges) and f-surp (capital income). The instruction is that f-surp needs to be split into wAges to entrepreneurs/self employed and capital income.

```{r}
#| warning: false
#| messAge: false
#| label: identify-incomes

pp_emstatus <- pp_emstatus |> 
  mutate(
    # We add accross three months for each source (and coalesce the NAs to 0)
    flab_job1 = 
      rowSums(
        across(starts_with("Q8_faqti_"), \(x) coalesce(x, 0))),
    flab_job2 = 
      rowSums(
        across(starts_with("Q14_faqti_"), \(x) coalesce(x, 0))),
    fsurp = 
      rowSums(
        across(starts_with("Q10_faqti_"), \(x) coalesce(x, 0)))
  ) |> 
  # We also add factor labels to Employment Status
  mutate(
    Q7 = factor(
      Q7,
      levels = c(1:6),
      labels = c(
        "Employee", "Employer", "Own Account (Non-Ag.)", 
        "Own Account (Ag.)", "Unpaid Worker", "Other"))
    ) |> 
  mutate(
    Q13 = factor(
      Q13,
      levels = c(1:6),
      labels = c(
        "Employee", "Employer", "Own Account (Non-Ag.)", 
        "Own Account (Ag.)", "Unpaid Worker", "Other"))
    )
```

And we add our labor market variables to our `pp_factor_descriptors` dataset.

```{r}
#| warning: false
#| messAge: false
#| label: join-factor-descriptors-EmStatus

pp_emstatus <- pp_emstatus |>
  select(-UID,-MemberNo) |> 
  left_join(pp_factor_descriptors, join_by(MemberId)) |> 
  relocate(c(UID, MemberNo, MemberId, QuartNo), .before = 1) |> 
  mutate(cohort = factor(case_when(
    Age >= 0  & Age < 16 ~ "P0015",
    Age >= 16 & Age < 30 ~ "P1629",
    Age >= 30 & Age < 40 ~ "P3039",
    Age >= 40 & Age < 50 ~ "P4049",
    Age >= 50 & Age < 60 ~ "P5059",
    Age >= 60  ~ "P60up"
  ), levels = c("P0015","P1629", "P3039", "P4049", "P5059", "P60up")))
```



# Own-account work in agriculture

In Georgia's Integrated Living Conditions Survey 2023 [@geostat_integrated_2023], own-account agriculture workers show no labor income by design. This follows new ILO recommendations, where agriculture income is tallied at the household level.

## Primary job

First identify the share of workers that have own account agriculture as primary job.

```{r}
test <- pp_emstatus |> 
  filter(!is.na(Q7)) |> 
  group_by(Q7) |> 
  summarize(
    People = sum(weight),
    LIncome1 = weighted.mean(flab_job1, weight, na.rm = T),
    SIncome1 = weighted.mean(fsurp, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)*100
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

## Secondary job

```{r}
test <- pp_emstatus |> 
  filter(!is.na(Q13)) |> 
  group_by(Q13) |> 
  summarize(
    People = sum(weight),
    LIncome2 = weighted.mean(flab_job2, weight, na.rm = T),
    SIncome1 = weighted.mean(fsurp, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)*100
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

## Own account work by activity

```{r}
test <- pp_emstatus |> 
  filter(!is.na(Q7) & SAM_job1 == "a-agri") |> 
  group_by(SAM_job1,Q7) |> 
  summarize(
    People = sum(weight),
    LIncome1 = weighted.mean(flab_job1, weight, na.rm = T),
    SIncome1 = weighted.mean(fsurp, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)*100
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

## Other income

Here we obtain the shares of different types of income for:

All Workers

```{r}
test <- pp_emstatus |> 
  filter(!is.na(Q7) ) |> 
  group_by(Quintile) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```



Employee

```{r}
test <- pp_emstatus |> 
  filter(Q7 == "Employee" & SAM_job1 == "a-agri") |> 
  group_by(Quintile) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Own Account (Ag.)

```{r}
test <- pp_emstatus |> 
  filter(Q7 == "Own Account (Ag.)" & SAM_job1 == "a-agri") |> 
  group_by(Quintile) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Unpaid Worker

```{r}
test <- pp_emstatus |> 
  filter(Q7 == "Unpaid Worker" & SAM_job1 == "a-agri") |> 
  group_by(Quintile) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```


## Age groups

All workers

```{r}

test <- pp_emstatus |> 
  filter(!is.na(Q7)) |> 
  group_by(cohort) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Employees

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Employee" & SAM_job1 == "a-agri") |> 
  group_by(cohort) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Own account

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Own Account (Ag.)" & SAM_job1 == "a-agri") |> 
  group_by(cohort) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Unpaid

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Unpaid Worker" & SAM_job1 == "a-agri") |> 
  group_by(cohort) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

## Gender

All workers

```{r}

test <- pp_emstatus |> 
  filter(!is.na(Q7)) |> 
  group_by(Gender) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Employees

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Employee" & SAM_job1 == "a-agri") |> 
  group_by(Gender) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Own account

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Own Account (Ag.)" & SAM_job1 == "a-agri") |> 
  group_by(Gender) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```

Unpaid

```{r}

test <- pp_emstatus |> 
  filter(Q7 == "Unpaid Worker" & SAM_job1 == "a-agri") |> 
  group_by(Gender) |> 
  mutate(
    shr_lab = coalesce(labinc,0.0001) / totalinc,
    shr_slf = coalesce(selfemp,0.0001) / totalinc,
    shr_agr = coalesce(aginc ,0.0001)/ totalinc,
    shr_ppt = coalesce(propertinc ,0.0001)/ totalinc,
    shr_ssp = coalesce(pensions,0.0001)/ totalinc,
    shr_rem = coalesce(remittances,0.0001)/ totalinc,
    shr_gft = coalesce(gifts,0.0001)/totalinc
  ) |> 
  summarize(
    People = sum(weight),
    avg_shr_lab = weighted.mean(shr_lab, weight, na.rm = T),
    avg_shr_slf = weighted.mean(shr_slf, weight, na.rm = T),
    avg_shr_agr = weighted.mean(shr_agr, weight, na.rm = T),
    avg_shr_ssp = weighted.mean(shr_ssp, weight, na.rm = T),
    avg_shr_rem = weighted.mean(shr_rem, weight, na.rm = T),
    avg_shr_gft = weighted.mean(shr_gft, weight, na.rm = T)
  ) |> 
  ungroup() |>
  mutate(
    Percent = People / sum(People, na.rm = T)
  ) |> 
  relocate(Percent, .after = People)

write.table(test, file = pipe("xclip -selection clipboard"), sep = "\t", row.names = FALSE)
```


# Save to Excel 


```{r}
# Define the file path
file_path <- "data/outputs/own_agriculture_data.xlsx"

# Check if the file exists
if (file.exists(file_path)) {
  # If the file exists, load the workbook
  wb <- wb_load(file_path)
} else {
  # If the file doesn't exist, create a new workbook
  wb <- wb_workbook()
}

# Print the sheet names in the workbook
sheet_names <- wb_get_sheet_names(wb)
sheet_names

# Define objects to add
objects_to_add <- c(
  "share_own_ag"
)

# Define sheets to add
sheets_to_add <- c(
  "Share Own Ag"
  )

for (sheet in sheets_to_add) {
  # Add content to the workbook
  if (! sheet %in% names(wb)) {
  # Add a new sheet
  wb_add_worksheet(wb, sheet)}
}

flexlsx::wb_add_flextable(
  wb,
  sheet = "Share Own Ag",
  share_own_ag,
  start_col = 1,
  start_row = 3
)

# for (i in seq_along(objects_to_add)){
#   writeData(
#   wb,
#   sheets_to_add[i], 
#   get(objects_to_add[i]), 
#   startRow = 5, 
#   startCol = 1, 
#   rowNames = FALSE)
# }

wb$save(file = "data/outputs/own_agriculture_data.xlsx", overwrite = T)
```