---
title: "SAM Labor Force Survey Data Request"
subtitle: "Georgia Country Climate and Development Report"
author:
  - name: "Renato Vargas"
    id: rv
    # email: renovargas@gmail.com
    affiliation: 
      - name: Consultant for The World Bank
  - name: "Noe Nicolas Reidt"
    id: nr
    # email: 
    affiliation: 
      - name: The World Bank
format: 
  html:
    toc: true
  # docx:
  #   toc: true
  #   number-sections: true
  #   highlight-style: arrow
  # pdf:
  #   # prefer-html: true
  #   toc: true
  #   number-sections: true
  #   colorlinks: true
editor_options: 
  chunk_output_type: console
---

# Introduction

For this data request, we summarize labor factor payments, by sex, occupation, employment status, and economic activity, according to a structure needed for the macroeconomic modeling.

## Objectives

1. Aggregate labor cost data according to the following dimensions:
    a.  Employment status.
    b.  Sex.
    c.  Occupation (according to ISCO-08).
    d.  Economic activity.
2. Additional data nodes:
    a. Number of workers.
    b. Labor cost.
    c. Number of hours (labor volume).
3. Output to Excel.

## Preliminaries

```{r}
#| warning: false
#| message: false
#| label: preliminaries

# Clean workspace
rm(list = ls())

# Georgia country ISO code
iso <- "GEO"

# Survey year
survey_year <- 2023

# Exchange rate USD per GEL
er <- 0.37

# Quarter of interest
quarter <- 107

# Years of interest for our macroeconomic scenario analysis
# analysis_years <- c(2030, 2050)
```

We will use the following libraries for this exercise.

```{r}
#| warning: false
#| message: false
#| label: libs

library(tidyverse)
library(haven)
library(readxl)
library(openxlsx)
library(gt)
```

## Datasets and correspondence tables

We use data from the 2023 Labor Force Survey and focus on quarter 1 (107).

```{r}
#| warning: false
#| message: false
#| label: datasets
#| lst-label: original-datasets

# Labor
pp_labor <- read_sav("data/lfs_2023/LFS_ECSTAT_ENG_2023.sav")

# Equivalence table
```

# Data transformation

Variables need to be ordered in a specific way. We do this with the use of factors. Fortunately, some of the heavy lifting has been done for us regarding occupations, but we still need to match economic activities to the SAM structure.

## Employment status

```{r}
# Job 1
pp_labor |> 
  rename(status1 = Status)
```

